---
title: 'Week 5 Coding: T-tests, Cohen''s d, one-way ANOVA'
author: "Travis Christy"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

First we are doing some exploring

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

# Quantile Quantile Plots
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

# Get descriptive statistics with summary table
chinstraps %>% 
  group_by(sex) %>% 
  summarise(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```
Re: Histogram - The distribution overall looks somewhat normal (bell shaped)
  
Re: QQ Plot - Linear distribution means they are close to normal (this is what very close to normal looks like)
  
  
## Get Vectors of body_mass_g for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

## T-Tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins? (Use two-sided t-test)

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_m, chinstraps_f)
```

You can see from the t-test the P-value allows us to reject null hypothesis
You can see from the cohen's d test, that there is a large effect size
  

One-sided t-test: Are male chinstrap penguins significantly larger (in body mass) than female chinstrap penguins?

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```












